Static Race Detection in OS Kernels by Mining Locking Rules

### 2023-07-21 1.introduction.tex
Update technique

(T1) \sys uses an {\em alias-aware rule mining method} to automatically 
deduce locking rules. We observe that an access to a data structure field is 
often protected by the lock in the same data structure as the accessed field. 
And this relationship between accessed field and the protecting lock can be 
effectively inferred from a special data structure named alias 
graph~\cite{Li:ASPLOS22, Kastrinis:CC18}. Specifically, fields in the same data 
structure have the same ancestor in the alias graph, and thus whether an 
accessed field and a specific lock are in the same data structure can be 
inferred by finding a common ancestor in the alias graph. Moreover, with 
benefits from precise field-sensitive alias relationships of alias graph, 
\sys can accurately extract the accessed field and its corresponding 
protecting lock. After obtaining the accessed field and protecting lock pairs, 
\sys calculates the proportion of field accesses protected by a specific 
lock in all field accesses. If the proportion is larger than a given threshold, 
the accessed field is determined to be protected by the protecting lock.

(T2) \sys uses a {\em lock-usage analysis} to filter out false positives 
caused by concurrency. After mining locking rules, \sys detects data 
races by checking whether a given data structure field access violates the 
locking rules. To reduce false negatives, \sys conservatively assumes 
every two functions can be executed concurrently, and thus can introduce false 
positives. However, we observe that each kernel module has an initialization 
phase, after which many functions can be called concurrently by other modules. 
When performing initialization, the kernel module serially initialized the lock 
and prepare other data for subsequent operations. And thus functions with the 
lock initialization and functions called by them tend not to be executed 
concurrently. Based on this observation, \sys extracts all functions 
reachable from the function with a lock initialization operation such as {\tt 
spin\_lock\_init()}, and suppose that this function can not be executed in 
parallel.

>>>>

(T1) \sys uses an {\em alias-aware rule mining method} to automatically 
deduce locking rules. We observe that an access to a data structure field is 
often protected by the lock in the same data structure as the accessed field. 
And this relationship between accessed field and the protecting lock can be 
effectively inferred from a special data structure named alias 
graph~\cite{Li:ASPLOS22, Kastrinis:CC18}. Specifically, fields in the same data 
structure have the same ancestor in the alias graph, and thus whether an 
accessed field and a specific lock are in the same data structure can be 
inferred by finding a common ancestor in the alias graph. Moreover, with 
benefits from precise field-sensitive alias relationships of alias graph, 
\sys can effectively extract the accessed field and its corresponding 
protecting lock. After obtaining the accessed field and protecting lock pairs, 
\sys calculates the proportion of field accesses protected by a specific 
lock in all field accesses. If the proportion is larger than a given threshold, 
the accessed field is determined to be protected by the protecting lock.

(T2) \sys uses a {\em lock-usage analysis} to filter out false positives 
validating concurrency of code paths. After mining locking rules, \sys detects 
data races by checking whether a given data structure field access violates the 
locking rules. To reduce false negatives, \sys conservatively assumes 
every two functions can be executed concurrently, and thus can introduce false 
positives. However, we observe that each kernel module has an initialization 
phase, after which many functions can be called concurrently by other modules. 
When performing initialization, the kernel module serially initialized the lock 
and prepare other data for subsequent operations. And thus functions with the 
lock initialization and functions called by them tend not to be executed 
concurrently. Based on this observation, \sys extracts all functions 
reachable from the function with a lock initialization operation such as {\tt 
spin\_lock\_init()}, and suppose that this function can not be executed in 
parallel.

### 2023-07-21 3.technique.tex
Update technique

% !TeX spellcheck = en_US
\section{Race Detection by Mining Locking Rules}
\label{sec_technique}

>>>>

% !TeX spellcheck = en_US
\section{Race Detection by Mining Locking Rules}
\label{sec_technique}
To address the above challenges, we propose three key techniques. For {\em C1}, 
we propose an {\em alias-aware rule mining method} to automatically deduce 
locking rules. For {\em C2}, we propose a {\em lock-usage analysis} to filter 
out false data races by validating concurrency of code paths. For {\em C3}, we 
propose a pattern-based estimation to extract harmful races that can trigger 
memory or logical bugs such as null-pointer dereference and data inconsistency. 
We introduce them as follows:

\subsection{Alias-Aware Rule Mining Method}
\label{subsec_rule_mining}
The relationship between variables and locks are not well documented in OS 
kernels, but it can be inferred from the kernel code. Specifically, a data 
structure field is often protected by the lock stored in the same data 
structure. And thus if a data structure field is accessed after acquiring a 
lock existing in the same data structure in most cases, the variable is likely 
to be protected by the lock. Whether a data structure field and the protecting 
lock exist in the same data structure can be determined though an alias 
graph~\cite{Li:ASPLOS22, Kastrinis:CC18} by finding their common ancestor. 
Based on this insight, we propose an {\em alias-aware rule mining method} to 
deduce locking rules automatically. Moreover, with benefits from precise 
field-sensitive alias relationships of alias graph, our alias-aware rule mining 
method can find data structure filed and its protecting lock effectively.

\PP{Alias Graph.} It is an important data structure to infer relationships 
between data structure field and its protecting lock in our analysis, so we 
introduce it and its update first.

An alias graph is a 2-tuple $\mathit{G = \left<N, E\right>}$, where 
$\mathit{N}$ is a set of nodes, and each node $\mathit{n}$ represents an alias 
set that points to one abstract object. $\mathit{E}$ is a set of labeled edges. 
Each edge is labeled with a data structure field or a dereference operator 
``$\mathit{*}$'', which represents how an abstract object is accessed.

An alias graph is updated by handling four types of instructions that  
change alias relationships: MOVE($\mathit{v_1 = v_2}$), STORE ($\mathit{*v_2 = 
v_1}$), LOAD ($\mathit{v_1 = *v_2}$) and GEP ({$\mathit{v_1 = \&v_1->f}$}). We 
exploit {$\mathit{n_x}$} to represent the node whose representing alias set 
includes $\mathit{v_x}$, and introduce how the four types of instructions 
update alias graphs. For a MOVE operation, $\mathit{v_1}$ is moved from 
$\mathit{n_1}$ to $\mathit{n_2}$. After this operation, $\mathit{v_1}$ and  
$\mathit{v_2}$ are represented by the same node, which indicates they become 
aliases. For a STORE operation, the existing outgoing edge from $\mathit{n_2}$ 
is dropped first, and then a new edge labeled with $\mathit{*}$ from 
$\mathit{n_2}$ to $\mathit{n_1}$ is inserted. After this operation, 
$\mathit{v_1}$ and $\mathit{*v_2}$ are represented by the same node, which 
indicates they become aliases. For a LOAD operation, the analysis first finds 
the destination node of the edge that comes from $\mathit{n_2}$ and is labeled 
with $\mathit{*}$, and then move $\mathit{v_1}$ to the destination node. And 
after this operation, $\mathit{v_1}$ and $\mathit{*v_2}$ are represented by the 
same node, which indicates they become aliases. GEP operation is similar to 
LOAD, expect that the edge is labeled with a data structure field $\mathit{f}$, 
instead of a dereference operator $\mathit{*}$.

\begin{figure}[htbp]
	\centering
	\includegraphics[width=0.9\linewidth]{figures/fig_alias_graph_demo.pdf}
	\figcaption{Example of alias graph.}
	\label{fig_alias_graph_demo}
\end{figure}

\noindent{\textbf{\em Example.}} Figure~\ref{fig_alias_graph_demo} shows a 
piece of driver-like source code and its alias graph. In this example, after an 
GEP (\&dev->lock) operation at Line 6, an edge labeled with {\tt lock} from 
node $\mathit{n_1}$ to node $\mathit{n_4}$ is inserted. Similarly, an edge 
labeled with {\tt flag} from node $\mathit{n_1}$ to node $\mathit{n_2}$ is 
inserted after Line 7. At last, an edge labeled with a dereference operator 
($\mathit{*}$) is inserted after the STORE (*fp = f) operation at Line 8. In 
this example, {\tt \&dev->flag} is represented by node $\mathit{n_2}$, and {\tt 
\&dev->lock} is represented by node $\mathit{n_4}$. The two nodes have a common 
ancestor $\mathit{n_1}$, and thus the accessed data structure field {\tt 
\&dev->flag} and the protecting lock {\tt \&dev->lock} can be inferred to exist 
in the same data structure (namely {\tt Dev}).
